Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-02-11 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd
starting to edit thermal tolerance section: insert new line per sentence

$\ddagger$ 1. Mau et al. 2018; 2. Coble and Cavaleri 2014; 3. Sack et al. 2006; 4. Chin and Sillett 2019; 5. Wyka et al. 2012; 6. Atherton et al. 2017; 7. Kenzo et al. 2015; 8. Kusi and Karasi 2020; 9. Dang et al. 1997; 10. Gebauer et al. 2015; 11. Marenco et al. 2017; 12. Kafuti et al. 2020; 13. Van Wittenberghe et al. 2012; 14. Zhang et al. 2019; 15. Weerasinghe et al. 2014; 16. Oldham et al. 2010; 17. Ichie et al. 2016; 18. Gregoriou et al. 2007; 19. Levizou et al. 2005; 20. Liakoura 1997; 21. Fauset et al. 2018; 22. Niinemets et al. 1998, 23. Ishida et al. 1998; 24. Millen and Clendon 1979; 25. Smith and Carter, 1988; 26. Hadley and Smith 1987; 28. Baltzer and Thomas 2005; 29. Coble et al. 2016; 30. Scartazza et al. 2016; 31. Duursma and Marshall, 2006; 32. Harley et al. 1996;  33. Hernandez et al. 2020; 34. Turnbull et al. 2003; 35. Chen et al. 2020; 36. van de Weg et al. 2012; 37. M.A Cavaleri et al. 2008;  38. Koniger et al. 1995; 39. Mastubara et al. 2009; 40. Harris and Medina 2013; 41. Hansen et al. 2001; 42. Poorter et al. 1995; 43. Coble et al. 2016; 44. Niinemets et al. 2004; 45. Poorter et al. 2000; 46. Zwieniecki et al. 2004; 47. Sack and Scoffoni, 2013; 48. Ball et al., 1988; 49. Taylor et al. 2021; 50. Niinemets et al. 2010; 51. Brooks et al. 1997 
```{r eval=FALSE, echo=FALSE}
# refs in this table currently in the nocite field at top of document 
```

@@

 "In Australia, along a vertical gradient within a very open canopy, Acacia papyrocarpa showed greater T50 and higher composite climate stress in the lower, sun-facing canopy than other canopy positions, correlating with low wind speed, greater radiation and Tair, and lower rates of heat dissipation, as indicated by longer leaf thermal time constants (Curtis et al., 2019). These patterns of higher heat tolerance in the warmest leaves are consistent with interspecific patterns across temperature gradients (O’Sullivan et al. 2018; Feeley et al., 2020; Slot et al., 2021a), and experienced maximum temperatures (Perez & Feeley, 2020). However, not enough studies have evaluated thermal tolerances across vertical gradients to make robust inferences [sorry, bit of a blabla sentence to wrap up]."

@@

$\ddagger$ 1. Kafuti et al. 2020; 2. Van Wittenberghe et al. 2012; 3. Roberts et al. 1990; 4. Dang et al. 1997; 5. Marenco et al. 2017; 6. Ambrose et al. 2015; 7. Zweifel et al. 2001; 8. Slot et al. 2019; 9. Hernandez et al. 2020; 10. Urban et al. 2007; 11. Carter and Cavaleri 2018; 12. Martin et al. 1999; 13. Mau et al. 2018;  14. Kosugi et al. 2012; 15. Niinemets et al. 2015; 16. Bachofen et al. 2020; 17. Hamerlynck and Knapp 1994; 18. Coble et al. 2017; 19. Wyka et al. 2012; 20. Rijkerse et al. 2000; 21. Ishida et al. 1999; 22. Weerasinghe et al. 2014; 23. Scartazza et al. 2016; 24. Miller et al. 2021; 25. Harris and Medina 2013; 26. Legner et al. 2014; 27. Kitao et al. 2012; 28. Fauset et al. 2018; 29. Rey-Sanchez et al. 2016; 30. Muller et al. 2021; 31. Curtis et al. 2019; 32. Mier et al. 2001; 33. Turnbull et al. 2003; 34. Araki et al. 2017; 35. Bolstad et al. 1999; 36. Kenzo et al. 2015; 37. Harley et al. 1996; 38. Xu and Griffin 2006; 39. Atherton et al. 2017; 40. Carter et al. 2021; 41. Sack et al. 2003; 42. Taylor et al. 2021; 43. Harley et al. 1997; 44. Niinemets and Sun, 2014; 45. Sharkey and Monson, 2014; 46. Simpraga et al. 2013

>>>>

$\ddagger$ 1. Mau et al. 2018; 2. Coble and Cavaleri 2014; 3. Sack et al. 2006; 4. Chin and Sillett 2019; 5. Wyka et al. 2012; 6. Atherton et al. 2017; 7. Kenzo et al. 2015; 8. Kusi and Karasi 2020; 9. Dang et al. 1997; 10. Gebauer et al. 2015; 11. Marenco et al. 2017; 12. Kafuti et al. 2020; 13. Van Wittenberghe et al. 2012; 14. Zhang et al. 2019; 15. Weerasinghe et al. 2014; 16. Oldham et al. 2010; 17. Ichie et al. 2016; 18. Gregoriou et al. 2007; 19. Levizou et al. 2005; 20. Liakoura 1997; 21. Fauset et al. 2018; 22. Niinemets et al. 1998, 23. Ishida et al. 1998; 24. Millen and Clendon 1979; 25. Smith and Carter, 1988; 26. Hadley and Smith 1987; 28. Baltzer and Thomas 2005; 29. Coble et al. 2016; 30. Scartazza et al. 2016; 31. Duursma and Marshall, 2006; 32. Harley et al. 1996;  33. Hernandez et al. 2020; 34. Turnbull et al. 2003; 35. Chen et al. 2020; 36. van de Weg et al. 2012; 37. M.A Cavaleri et al. 2008;  38. Koniger et al. 1995; 39. Mastubara et al. 2009; 40. Harris and Medina 2013; 41. Hansen et al. 2001; 42. Poorter et al. 1995; 43. Coble et al. 2016; 44. Niinemets et al. 2004; 45. Poorter et al. 2000; 46. Zwieniecki et al. 2004; 47. Sack and Scoffoni, 2013; 48. Ball et al., 1988; 49. Taylor et al. 2021; 50. Niinemets et al. 2010; 51. Brooks et al. 1997; 52. Curtis et al. 2019; 53. Slot et al. 2019; 54. Hamerlynck and Knapp 1994.
```{r eval=FALSE, echo=FALSE}
# refs in this table currently in the nocite field at top of document 
```

@@

In Australia, along a vertical gradient within a very open canopy, Acacia papyrocarpa showed greater T50 and higher composite climate stress in the lower, sun-facing canopy than other canopy positions, correlating with low wind speed, greater radiation and Tair, and lower rates of heat dissipation, as indicated by longer leaf thermal time constants (Curtis et al., 2019). 
These patterns of higher heat tolerance in the warmest leaves are consistent with interspecific patterns across temperature gradients (O’Sullivan et al. 2018; Feeley et al., 2020; Slot et al., 2021a), and experienced maximum temperatures (Perez & Feeley, 2020). 
However, not enough studies have evaluated thermal tolerances across vertical gradients to make robust inferences.

@@

$\ddagger$ 1. Kafuti et al. 2020; 2. Van Wittenberghe et al. 2012; 3. Roberts et al. 1990; 4. Dang et al. 1997; 5. Marenco et al. 2017; 6. Ambrose et al. 2015; 7. Zweifel et al. 2001; 8. Slot et al. 2019; 9. Hernandez et al. 2020; 10. Urban et al. 2007; 11. Carter and Cavaleri 2018; 12. Martin et al. 1999; 13. Mau et al. 2018;  14. Kosugi et al. 2012; 15. Niinemets et al. 2015; 16. Bachofen et al. 2020; 17. Hamerlynck and Knapp 1994; 18. Coble et al. 2017; 19. Wyka et al. 2012; 20. Rijkerse et al. 2000; 21. Ishida et al. 1999; 22. Weerasinghe et al. 2014; 23. Scartazza et al. 2016; 24. Miller et al. 2021; 25. Harris and Medina 2013; 26. Legner et al. 2014; 27. Kitao et al. 2012; 28. Fauset et al. 2018; 29. Rey-Sanchez et al. 2016; 30. Muller et al. 2021; 31. Mier et al. 2001; 32. Turnbull et al. 2003; 33. Araki et al. 2017; 34. Bolstad et al. 1999; 35. Kenzo et al. 2015; 36. Harley et al. 1996; 37. Xu and Griffin 2006; 38. Atherton et al. 2017; 39. Carter et al. 2021; 40. Sack et al. 2003; 41. Taylor et al. 2021; 42. Harley et al. 1997; 43. Niinemets and Sun, 2014; 44. Sharkey and Monson, 2014; 45. Simpraga et al. 2013; 46. Slot et al. 2021
